### 2025-05-14 paper/tables/vocab_overlap.tex

\begin{table}[htbp]
\centering
\scriptsize
\setlength{\tabcolsep}{4pt}
\begin{tabular}{l|cc}
\toprule
& \multicolumn{2}{c}{\textbf{Test Vocabulary Coverage}} \\
\textbf{Task} & \textbf{OOV Tokens} & \textbf{OOV\%} \\
\midrule
SST2 & 34 & 0.42\% \\
MRPC & 465 & 13.04\% \\
COLA & 169 & 8.74\% \\
STSB & 448 & 13.70\% \\
QQP & 191 & 1.03\% \\
MNLI & 31 & 0.14\% \\
QNLI & 83 & 0.41\% \\
RTE & 509 & 12.12\% \\
\bottomrule
\end{tabular}
\caption{Train-test vocabulary overlap statistics. \textbf{OOV Tokens}: number of unique tokens in the test set that don't appear in the training set; \textbf{OOV\%}: percentage of test vocabulary not found in train. Lower percentages indicate better vocabulary coverage, which is beneficial for pruning.}
\label{tab:vocab_overlap}
\end{table}