### 2025-05-23 paper/tables/benchmark_results.tex
update tables

\begin{table}[htb]
\centering
\scriptsize
\setlength{\tabcolsep}{3pt}
\begin{tabular}{lrrr}
\toprule
\textbf{Metric} & \textbf{Base} & \textbf{Pruned} & \textbf{Impr.(\%)} \\
\midrule
Params (M) & 149.61 & 119.69 & 20.00 \\
Embed. Params (M) & 38.68 & 8.76 & 77.34 \\
Storage (MB) & 570.72 & 456.59 & 20.00 \\
GPU Mem (MB) & 823.06 & 700.99 & 14.83 \\
Infer. Time (ms) & 18.73 & 18.98 & -1.33 \\
\bottomrule
\end{tabular}
% \caption{Benchmark results comparing base ModernBERT with pruned variant.}
% \caption{Benchmark results comparing base ModernBERT with pruned variant during inference on an RTX 4090.}
\caption{Benchmark results comparing base ModernBERT with pruned variant (TF-IDF + OOV) during inference on an RTX 4090. Our method achieves 20\% reduction in total parameters (77.34\% of embedding parameters) and a 14.83\% reduction in GPU memory with negligible impact on inference time.}

\label{tab:benchmark_results}
\end{table}