### 2025-05-23 paper/tables/token_statistics.tex

\begin{table}[htbp]
\centering
\scriptsize
\setlength{\tabcolsep}{3pt}
\begin{tabular}{l|rr|rr|rr|r}
\toprule
\textbf{Task} & \multicolumn{2}{c|}{\textbf{Unique Tokens}} & \multicolumn{2}{c|}{\textbf{Vocab Cov. (\%)}} & \multicolumn{2}{c|}{\textbf{Top 20\%}} & \textbf{OOV} \\
    & \textbf{Train} & \textbf{Test} & \textbf{Train} & \textbf{Test} & \textbf{Train} & \textbf{Test} & \textbf{\%} \\
\midrule
COLA & 5,416 & 1,934 & 17.74 & 6.34 & 85.12 & 74.68 & 8.74 \\
MNLI & 25,793 & 22,664 & 84.51 & 74.25 & 90.69 & 89.64 & 0.14 \\
MRPC & 11,096 & 3,567 & 36.35 & 11.69 & 83.17 & 71.80 & 13.04 \\
QNLI & 26,176 & 20,360 & 85.76 & 66.71 & 88.31 & 85.79 & 0.41 \\
QQP & 25,486 & 18,534 & 83.50 & 60.72 & 94.02 & 91.79 & 1.03 \\
RTE & 13,354 & 4,198 & 43.75 & 13.75 & 81.77 & 70.75 & 12.12 \\
SST2 & 11,536 & 8,084 & 37.80 & 26.49 & 84.23 & 80.53 & 0.42 \\
STSB & 10,346 & 3,271 & 33.90 & 10.72 & 82.66 & 71.30 & 13.70 \\
\bottomrule
\end{tabular}
\caption{Token statistics across GLUE tasks. Vocab Cov.: percentage of model vocabulary present in data. Top 20\%: token coverage by most frequent 20\% of vocabulary. OOV\%: test tokens not in training.}

\label{tab:token_statistics}
\end{table}